---
title: "NC Genes ATAC-Seq Signal"
output: html_notebook
---

## Running Annotation on diffbind
```{r Annotated peakset}
library(ChIPseeker)
library(RMariaDB)
library(org.Gg.eg.db)
library(clusterProfiler)
library(ReactomePA)
library(tidyr)
library(rtracklayer)
library(GenomicFeatures)
library(DiffBind)
library(pheatmap)
library(stringr)
library(dplyr)
library(ggplot2)

#Annotating all ATAC peaks
samples_ATAC <- readRDS(file = "./Exports/samples_ATAC_400bp_2.RDS")
ATAC_all_peaks = readPeakFile("Imports/all_ATAC_400bp_fixed.bed")
TxDb_galGal6 <- makeTxDbFromEnsembl(organism = "Gallus gallus", release = 101)

ATAC_all_peaks_anno = annotatePeak(ATAC_all_peaks, tssRegion = c(-1000,1000), TxDb = TxDb_galGal6, annoDb = "org.Gg.eg.db")
ATAC_all_peaks_anno.df = data.frame(ATAC_all_peaks_anno)

GRN_Genes <- read.csv(file = "./Imports/GRN_Genes.csv", header = F)
colnames(GRN_Genes) <- c("Symbol","ENSEMBL")
#combining with counts peaks
#rerun dba.peakset for all ATAC bam files as done previously

# TMM Normalized counts
normCounts <- dba.peakset(samples_ATAC, bRetrieve=TRUE, DataType=DBA_DATA_FRAME)
normCounts$START = normCounts$START + 1

colnames(normCounts)[1:3] = c("seqnames", "start", "end")

ATAC_all_score_anno = merge(normCounts, ATAC_all_peaks_anno.df, by = c("seqnames", "start", "end"))

ATAC_all_score_anno$start = ATAC_all_score_anno$start - 1
#Determining ATAC peaks around NC genes
NC_peaks = subset(ATAC_all_score_anno, ATAC_all_score_anno$geneId %in% GRN_Genes$ENSEMBL)


NC_peaks$number = c(1:461)
NC_peaks_score = data.frame(NC_peaks[,4:17],
                            SYMBOL = NC_peaks$SYMBOL,
                            number = NC_peaks$number,
                            distance = round(NC_peaks$distanceToTSS/1000, digits = 1)
                            )

rownames(NC_peaks_score) <- paste0(NC_peaks_score$SYMBOL,"_",NC_peaks_score$number,"_",NC_peaks_score$distance)

boxplot(abs(NC_peaks_score$distance))

NC_peaks_avg <- data.frame(HH6 = rowMeans(NC_peaks_score[c('HH6_1', 'HH6_2')], na.rm=TRUE),
                         HH8 = rowMeans(NC_peaks_score[c('HH8_1', 'HH8_2')], na.rm=TRUE),
                         HH10 = rowMeans(NC_peaks_score[c('HH10_1', 'HH10_2')], na.rm=TRUE),
                         HH12 = rowMeans(NC_peaks_score[c('HH12_1', 'HH12_2')], na.rm=TRUE),
                         HH14 = rowMeans(NC_peaks_score[c('HH14_1', 'HH14_2')], na.rm=TRUE),
                         HH16 = rowMeans(NC_peaks_score[c('HH16_2', 'HH16_1')], na.rm=TRUE),
                         HH18 = rowMeans(NC_peaks_score[c('HH18_1', 'HH18_2')], na.rm=TRUE))


pheatmap(NC_peaks_avg,
         cluster_cols = F,
         scale = "row",
         show_rownames = F, 
         cutree_rows = 6, 
         fontsize_row = 6,
         border_color = NA,
         color = colorRampPalette(colors = c("#0797B3","#FFFFFF","#4FA14E"))(250),
         width = 4.5,
         height = 10, filename = "./Exports/NC_GRN_ATAC-Peaks.pdf")



pheatmap(NC_peaks_avg[abs(NC_peaks_score$distance) < 75,],
         cluster_cols = F,
         scale = "row",
         show_rownames = F, 
         cutree_rows = 6, 
         fontsize_row = 6,
         border_color = NA,
         color = colorRampPalette(colors = c("#0797B3","#FFFFFF","#4FA14E"))(250),
         width = 4.5,
         height = 10, filename = "./Exports/NC_GRN_ATAC-Peaks_lessthan75kb.pdf")


# Rewrite early module
early_genes <- c("MSX1", "PAX7", "SALL1","OTX2","LIN28A","NANOG",
                 "POU4F3", "SP5","DMBX1","GATA3","HAND2","TFAP2C",
                 "MSX2", "ZIC3", "ZIC1", "GBX2", "PAX3", "PRDM1",
                 "LHX5","Pou5f3")

early_genes <- c("MSX1", "PAX7","LIN28A","NANOG",
                 "SP5","DMBX1","GATA3","ZIC3", "ZIC1", "GBX2", "PAX3",
                 "Pou5f3")

logic <- str_split(rownames(NC_peaks_avg), pattern = "_", simplify = T)[,1] %in% early_genes
early_module_avg = NC_peaks_avg[logic,]
early_module_avg$Enhancer <- rownames(early_module_avg)
early_melt <- melt(early_module_avg, id.vars = c("Enhancer"), measure.vars = c(1,2,3,4,5,6,7))

early_melt <- early_melt %>% group_by(Enhancer) %>% mutate(scaled_val = scale(value))

lin <- lm(scaled_val ~ Enhancer, data = early_melt)
lin$qr

early_melt$chopped_time <- str_split(early_melt$variable, pattern = "HH", simplify = T)[,2]

# Throw out poor correlations?
early_melt <- early_melt %>% group_by(Enhancer) %>% mutate(correlation = cor(as.numeric(chopped_time), scaled_val))

early_melt$direction <- ifelse(early_melt$correlation > 0.5,"Positive",
                               ifelse(early_melt$correlation < -0.5, "Negative","Poor Correlation"))

ggplot(early_melt, aes(variable, scaled_val)) + 
  geom_point(aes(group = Enhancer)) + 
  geom_line(aes(group = Enhancer, color = direction)) + 
  geom_smooth(method="lm") + 
  facet_wrap(facets = vars(direction), ncol = 1) +
  theme()


good_early_enhancers <- early_melt[early_melt$direction == "Negative",]
unique(good_early_enhancers$Enhancer)

pheatmap(early_module_avg[unique(good_early_enhancers$Enhancer),1:7],
         cluster_cols = F,
         scale = "row",
         show_rownames = T,
         fontsize_row = 6,
         border_color = NA,
         color = colorRampPalette(colors = c("#0797B3","#FFFFFF","#4FA14E"))(250),
         clustering_method = "complete",
         width = 5,
         height = 12, filename = "./Exports/Early_NC_Module_ATAC-Peaks.pdf")


# Let's see if we can get good middle enhancers by fitting a quadratic model?

summary(test)$adj.r.squared
early_melt2 <- early_melt %>% mutate(quad = summary(lm(scaled_val ~ poly(as.numeric(chopped_time), degree = 2)))$adj.r.squared)

early_melt2$quad_fit <- ifelse(early_melt2$direction == "Poor Correlation" & early_melt2$quad > 0.5,
                               "Quadratic Fit", early_melt2$direction)

ggplot(early_melt2, aes(variable, scaled_val)) + 
  geom_point(aes(group = Enhancer)) + 
  geom_line(aes(group = Enhancer, color = quad_fit)) + 
  geom_smooth(method="lm") + 
  facet_wrap(facets = vars(quad_fit), ncol = 1) +
  theme()

middle_module = c("SNAI2","FOXD3","CSRNP1","SOX1O","RXRG","MYC","ZEB2","MYCN","ETS1","MEF2C","SOX5","TFAP2B","LMO4","COL2A1","SOX9","TWIST2","TFAP2A","SLIT3")
logic <- str_split(rownames(NC_peaks_avg), pattern = "_", simplify = T)[,1] %in% middle_module
middle_module_avg = NC_peaks_avg[logic,]

pheatmap(middle_module_avg, cluster_cols = F, scale = "row",
         show_rownames = T, fontsize_row = 8, cellheight = 10)

Late_module = c("MITF","TYR","TH","TWIST2","PHOX2B","MEF2C","DCT","EBF1")
logic <- str_split(rownames(NC_peaks_avg), pattern = "_", simplify = T)[,1] %in% Late_module
Late_module_avg = NC_peaks_avg[logic,]

pheatmap(Late_module_avg, cluster_cols = F, scale = "row",
         show_rownames = T, fontsize_row = 8, cellheight = 10)
```

This is looking pretty good for the previously established GRN genes, but what about TF's that we identified from our RNA-Seq analysis? It may help to link the datasets together better.

Additionally, maybe throw in a promoter overlap and RNA-Seq correlation plot?
Could compare Log2 ATAC Counts vs Log2 TPM RNA-Seq counts

Summary of last time. 
pearson correlation from tpm and rpm of RNA-Seq and ATAC-Seq is between 40-60% from what I've seen. This may differ at timepoints.
```{r}
setwd(dir = "/data/Debbie/Sox2_Oct4_project/exports/Fig1/BAM")
rawCounts <- dba.count(samples_ATAC, peaks = ATAC_all_peaks,
                       score = DBA_SCORE_READS)

# Raw counts
rawCounts_df <- dba.peakset(rawCounts, bRetrieve=TRUE, DataType=DBA_DATA_FRAME)
colnames(rawCounts_df)[1:3] = c("seqnames", "start", "end")

ATAC_Counts_Anno = merge(rawCounts_df, ATAC_all_peaks_anno.df, by = c("seqnames", "start", "end"))
```

```{r}
# Get peak counts over promoters
gene_promoters <- trim(promoters(TxDb_galGal6, upstream = 1000, downstream = 1000))

prom_peaks <- ATAC_Counts_Anno[ATAC_all_peaks %over% gene_promoters,]

prom_peak_score = data.frame(prom_peaks[,4:17],
                            SYMBOL = prom_peaks$SYMBOL,
                            geneID = prom_peaks$geneId,
                            number = c(1:length(prom_peaks$geneId)),
                            distance = round(prom_peaks$distanceToTSS/1000, digits = 1)
                            )

rownames(prom_peak_score) <- paste0(prom_peak_score$geneID,"_",prom_peak_score$number,
                                    "_", prom_peak_score$distance)

counts <- prom_peak_score[,1:14]
counts[is.na(counts)] <- 0
prom_peak_score[,1:14] <- counts

cor(prom_peak_score$HH6_1, prom_peak_score$HH6_2)

ggplot(prom_peak_score, aes(HH6_1, HH6_2)) + geom_point() + 
  geom_smooth(method = "lm", show.legend = T) + lims(x = c(0, 1000), y = c(0,1000))

prom_peak_avg <- data.frame(HH6 = rowMeans(prom_peak_score[c('HH6_1', 'HH6_2')], na.rm=TRUE),
                         HH8 = rowMeans(prom_peak_score[c('HH8_1', 'HH8_2')], na.rm=TRUE),
                         HH10 = rowMeans(prom_peak_score[c('HH10_1', 'HH10_2')], na.rm=TRUE),
                         HH12 = rowMeans(prom_peak_score[c('HH12_1', 'HH12_2')], na.rm=TRUE),
                         HH14 = rowMeans(prom_peak_score[c('HH14_1', 'HH14_2')], na.rm=TRUE),
                         HH16 = rowMeans(prom_peak_score[c('HH16_2', 'HH16_1')], na.rm=TRUE),
                         geneID = prom_peaks$geneId)

# Get raw counts for genes.
raw_count_genes <- read.csv(file = "~/local_git/NC_Timecourse/RNA-Seq/Exports/Raw_RNA-Seq_Counts.csv", row.names = 1)
raw_count_genes$Chick_ENSEMBL <- rownames(raw_count_genes)
# Remove peaks that don't have a gene from RNA-Seq.

prom_peak_avg_filt <- prom_peak_avg[prom_peak_score$geneID %in% raw_count_genes$Chick_ENSEMBL,]

# Collapse (mean) values for each gene

prom_peak_avg_filt_2 <- prom_peak_avg_filt %>% 
  group_by(geneID) %>% 
  summarise(HH6_av = mean(HH6),
            HH8_av = mean(HH8),
            HH10_av = mean(HH10),
            HH12_av = mean(HH12),
            HH14_av = mean(HH14),
            HH16_av = mean(HH16))


raw_count_genes_filt <- subset(raw_count_genes, raw_count_genes$Chick_ENSEMBL %in% prom_peak_avg_filt_2$geneID)

raw_count_genes_filt_tpm <- raw_count_genes_filt[,1:42]/(colSums(raw_count_genes_filt[,1:42])/1000000)
raw_count_genes_filt_tpm <- raw_count_genes_filt_tpm[order(rownames(raw_count_genes_filt_tpm)),]
# Get transcript lengths for these genes

genes <- genes(TxDb_galGal6)
ens_to_length <- data.frame(ENSEMBL = genes$gene_id,
                            width = genes@ranges@width)

raw_count_genes_filt$length <- qdapTools::lookup(terms = rownames(raw_count_genes_filt), ens_to_length)

tmm_genes_filt <- as.data.frame(NOISeq::tmm(raw_count_genes_filt[,1:42],
            long = raw_count_genes_filt$length,
            lc = 0, k = 0, refColumn = 1,
            logratioTrim = 0.3, sumTrim = 0.05, doWeighting = TRUE, Acutoff = -1e+10))

tmm_genes_filt <- tmm_genes_filt[order(rownames(tmm_genes_filt)),]

# Construct an x/y for every timepoint. Probably only need to show one but check them all? Maybe trend?
cor(prom_peak_avg_filt_2$HH10_av, raw_count_genes_filt_tpm$HH10_mGFP_2_GTAGAG)

hh6_comparison <- data.frame(rpm_atac = prom_peak_avg_filt_2$HH6_av,
                             tpm_rna = raw_count_genes_filt_tpm$HH6v3_pos_CGATGT)

cor(hh6_comparison$rpm_atac, hh6_comparison$tpm_rna)

ggplot(hh6_comparison, aes(rpm_atac, tpm_rna)) + geom_point() + 
  geom_smooth(method = "lm", show.legend = T) +
  lims(x = c(0, 600), y = c(0,1000))

# Try filtering.
# on ATAC
upper <- quantile(hh6_comparison$rpm_atac, 0.98)
lower <- 1
hh6_comparison_filt <- hh6_comparison[hh6_comparison$rpm_atac < upper,]
hh6_comparison_filt <- hh6_comparison[hh6_comparison$rpm_atac > lower,]

upper_rna <- quantile(hh6_comparison$tpm_rna, 0.98)
lower_rna <- 1
hh6_comparison_filt <- hh6_comparison[hh6_comparison$tpm_rna < upper_rna,]
hh6_comparison_filt <- hh6_comparison[hh6_comparison$tpm_rna > lower_rna,]
cor(hh6_comparison_filt$rpm_atac, hh6_comparison_filt$tpm_rna)

ggplot(hh6_comparison_filt, aes(rpm_atac, tpm_rna)) + geom_point() + 
  geom_smooth(method = "lm", show.legend = T) +
  lims(x = c(0, 600), y = c(0,1000))

```

## Importing diffTF results
I used (with great difficulty) the program diffTF to perform integration of our ATAC-Seq and RNA-Seq data. This program scans peaks for TFBS's, and then associates the accessibility of those sites in aggregate to the RNA-Seq expression levels of eligible genes. Thus, it is able to classify each transcription factor as an activator, repressor, or undetermined. Undetermined genes may have both activities, or no real specific activity.

```{r}
library(readr)
library(qdapTools)
output_global_TFs <- read_csv("Imports/output.global.TFs.csv")
output_global_TFs_orig <- read_csv("Imports/output.global.TFs.orig.csv")
all_timepoints_summary <- read_delim("Imports/all.timepoints.summary.tsv", 
                                     "\t", escape_double = FALSE, trim_ws = TRUE)
trans_table <- read_delim(file = "Imports/translationTable_gg6.csv", delim = " ")
# https://www.grnpedia.org/trrust/
trrust_rawdata_human <- read_delim("Imports/trrust_rawdata.human.tsv", 
                                   "\t", escape_double = FALSE, col_names = FALSE, 
                                   trim_ws = TRUE)
ddsTC_Condition_WE_vc_NC_All <- read_csv("../RNA-Seq/Exports/ddsTC_Condition_WE_vc_NC_All.csv")

all_timepoints_summary$Symbol <- lookup(all_timepoints_summary$TF, as.data.frame(trans_table[,c(3,1)]))
output_global_TFs$Symbol <- lookup(output_global_TFs$TF, as.data.frame(trans_table[,c(3,1)]))
output_global_TFs_orig$Symbol <- lookup(output_global_TFs_orig$TF, as.data.frame(trans_table[,c(3,1)]))

summary(as.factor(all_timepoints_summary$classification_q0.05_final))

activators <- all_timepoints_summary[all_timepoints_summary$classification_q0.05_final == "activator",]
repressors <- all_timepoints_summary[all_timepoints_summary$classification_q0.05_final == "repressor",]


represented <- all_timepoints_summary$Symbol[all_timepoints_summary$Symbol %in% trrust_rawdata_human$X1]
not_represented <- all_timepoints_summary$Symbol[!all_timepoints_summary$Symbol %in% trrust_rawdata_human$X1]

report <- data.frame()

for (i in seq(1,length(represented))) {
  db_sub <- subset(trrust_rawdata_human, trrust_rawdata_human$X1 == represented[i])
  act_count <- length(db_sub[db_sub$X3 == "Activation",4]$X4)
  rep_count <- length(db_sub[db_sub$X3 == "Repression",4]$X4)
  unkn_count <- length(db_sub[db_sub$X3 == "Unknown",4]$X4)
  report <- rbind(report, data.frame(represented[i], act_count, rep_count, unkn_count))
}

timepoints_by_symb <- as.data.frame(all_timepoints_summary[,c("Symbol","classification_q0.05_final")])
timepoints_by_symb <- timepoints_by_symb[!is.na(timepoints_by_symb$Symbol),]
timepoints_by_symb <- timepoints_by_symb[!duplicated(timepoints_by_symb$Symbol),]
report$NC_timecourse_call <- qdapTools::lookup(report$represented.i., timepoints_by_symb)
report$conflict <- ifelse(report$NC_timecourse_call == "activator" & report$rep_count > report$act_count, 
                          "conflict", "expected")
```

##  Let's plot some things.
Timecourse RNA-Seq expression of the activators and repressors
Next to the chromvar scores for each time.
```{r}
library(pheatmap)
library(stringr)
library(ggplot2)
library(heatmaply)
library(RColorBrewer)


Annotated_Rlog_Positive_Averages <- read_csv("../RNA-Seq/Exports/Annotated_Rlog_Positive_Averages.csv")

Rlog_counts_avergae_All_ATAC_Peaks <- read.delim("~/local_git/NC_Timecourse/ATAC-Seq/Exports/Rlog_counts_avergae_All_ATAC_Peaks")



chromVAR_Z_Score <- as.data.frame(read_csv("Exports/chromVAR_Z-Score.csv"))
rownames(chromVAR_Z_Score) <- chromVAR_Z_Score$X1
chromVAR_Z_Score$X1 <- NULL

names <- data.frame(ID =
str_split(rownames(chromVAR_Z_Score), "_", simplify = T)[,1],
name = 
str_split(rownames(chromVAR_Z_Score), "_", simplify = T)[,2]
)
all_timepoints_summary$TF_and_Symbol <- paste0(all_timepoints_summary$TF, "_",qdapTools::lookup(all_timepoints_summary$TF, names))
all_timepoints_summary[all_timepoints_summary$classification_q0.1 == "activator" | all_timepoints_summary$classification_q0.1 == "repressor",]$TF

timepoint_sub_1 <- all_timepoints_summary[all_timepoints_summary$classification_q0.1 == "activator",]
timepoint_sub_2 <- all_timepoints_summary[all_timepoints_summary$classification_q0.1 == "repressor",]

act <- chromVAR_Z_Score[rownames(chromVAR_Z_Score) %in% timepoint_sub_1$TF_and_Symbol,]
rep <- chromVAR_Z_Score[rownames(chromVAR_Z_Score) %in% timepoint_sub_2$TF_and_Symbol,]

pheatmap(act, cluster_cols = F, scale = "row",
         border_color = NA, show_rownames = F,
         color = brewer.pal(250, "PRGn"),
         main = "Activators - chromVAR")
cowplot::plot_grid(act_hm, rep_hm) 
```

Let's try making a plot with the impulse DE results. We have Rlog counts and a list of genes and p values.

```{r}
library(data.table)
Impulse_Results <- read_csv("../RNA-Seq/Exports/Impulse_Results.csv")

fit = get_lsModelFits(obj=impulse_obj)
params_list = list()
for (gene in names(fit$case)){
  fit_params = data.table(t(fit$case[[gene]]$lsImpulseFit$vecImpulseParam))
  fit_params[,ENSEMBL := gene]
  params_list[[gene]] = fit_params
}
all_params = rbindlist(params_list)

#Return dfFit as well as the plot.
custom_plotGenes <- edit(plotGenes)


res_1 <- custom_plotGenes(
  vecGeneIDs       = NULL,
  scaNTopIDs       = NULL,
  objectImpulseDE2 = impulse_obj,
  boolCaseCtrl     = T,
  boolSimplePlot   = T,
  dirOut           = NULL,
  strFileName      = NULL,
  vecRefPval       = NULL, 
  strNameRefMethod = NULL)

vecTimePointsFit <- seq(
            min(get_dfAnnotationProc(obj=impulse_obj)$Time), 
                                max(get_dfAnnotationProc(obj=impulse_obj)$Time), 
                                length.out = 100)


```

So theoretically, I can use cor() to measure the similarity of two lines aka lines of best fit from both my Impulse results and ATAC-Seq peaks. Maybe I can use this to find peaks with similar dynamics as the gene in a more robust way.
```{r}
# Here's an idea, just run ImpulseDE2 on the timecourse ATAC-Seq data, then correlate each gene's peaks and find ones that match the RNA-Seq model.
#The annotation data has to be modified for this tool.

samples_ATAC2 <- dba.count(samples_ATAC, peaks=NULL, score=DBA_SCORE_READS)
counts <- dba.peakset(samples_ATAC2, bRetrieve=TRUE, DataType=DBA_DATA_FRAME)

rownames(counts) <- paste0(counts$CHR,"-", counts$START, "-", counts$END)
counts <- as.matrix(counts[,4:17])
counts <- counts[,1:12]
anno2 <- data.frame(Sample = colnames(counts),
                    Condition = "case",
                    Time = c(6,6,8,8,10,10,12,12,14,14,16,16),
                    Batch = "B_NULL")


impulse_obj2 <- runImpulseDE2(
  matCountData = counts,
  dfAnnotation = anno2,
  boolCaseCtrl = F,
  vecConfounders = NULL,
  boolIdentifyTransients = T,
  scaNProc = 16)

impulse_obj2$dfImpulseDE2Results

head(impulse_obj2$dfImpulseDE2Results)
res2 <- impulse_obj2$dfImpulseDE2Results

saveRDS(impulse_obj2, file = "Exports/ATAC-Seq-Impulse.RDS")
lsgplotsGenes2 <- plotGenes(
  vecGeneIDs       = NULL,
  scaNTopIDs       = 10,
  objectImpulseDE2 = impulse_obj2,
  boolCaseCtrl     = F,
  boolSimplePlot   = T,
  dirOut           = NULL,
  strFileName      = NULL,
  vecRefPval       = NULL, 
  strNameRefMethod = NULL)

lsgplotsGenes2[[1]]

```
1. Retrieve gene impulse model
2. For every peak for that gene, find the correlation
3. Get a set of random peaks that are not for that gene and find the average correlation for unrelated peaks. Did this, it is 0! 
4. Report, maybe determine some sort of cutoff?
```{r}
library(BiocParallel)
library(qdapTools)
library(ImpulseDE2)
register(BPPARAM = MulticoreParam(progressbar = T))
impulse_obj <- readRDS("../RNA-Seq/Exports/impulse_obj.RDS")
impulse_obj2 <- readRDS("Exports/ATAC-Seq-Impulse.RDS")
ATAC_all_peaks_anno.df <- readRDS("Exports/ATAC_all_peaks_anno.RDS")

#Return dfFit instead of plot and comment print message.
getDfFit <- edit(plotGenes)

compareGeneToPeaks <-
  function(gene = "") {
    # Get the RNA-Seq fit
    gene_fit <-
      getDfFit(
        vecGeneIDs = gene,
        objectImpulseDE2 = impulse_obj,
        boolCaseCtrl = T,
        boolSimplePlot = T
      )
    gene_fit <- gene_fit[gene_fit$Condition == "case", 2]
    
    # Filter the peak list for peaks associated with that gene.
    peaks_of_interest <-
      ATAC_all_peaks_anno.df[ATAC_all_peaks_anno.df$geneId %in% gene, "PeakID"]
    
    # Get the ATAC-Seq fit
    peak_res <- list()
    peak_dist <- list()
    peak_anno <- list()
    
    # Make sure it's in the ATAC-Seq obj.
    peaks_of_interest <-
      peaks_of_interest[peaks_of_interest %in% impulse_obj2@vecAllIDs]
    
    # If NONE of the peaks are in the object, return NA.
    
    if (!length(peaks_of_interest)) {
      return(data.frame(spearman = NA,
                      gene = gene,
                      distancetoTSS = NA,
                      annotation = NA))
    }
    
    for (peak in peaks_of_interest) {
      peak_fit <-
        getDfFit(
          vecGeneIDs = peak,
          objectImpulseDE2 = impulse_obj2,
          boolCaseCtrl = F,
          boolSimplePlot = T
        )
      peak_fit <- peak_fit[, 2]
      #spearman correlation
      peak_res[peak] <- cor(gene_fit, peak_fit, method = "spearman")
      peak_dist[peak] <- ATAC_all_peaks_anno.df[ATAC_all_peaks_anno.df$PeakID == peak, "distanceToTSS"]
      peak_anno[peak] <- ATAC_all_peaks_anno.df[ATAC_all_peaks_anno.df$PeakID == peak, "annotation"]
      
    }
    
    return(data.frame(spearman = unlist(peak_res),
                      gene = gene,
                      distancetoTSS = unlist(peak_dist),
                      annotation = unlist(peak_anno)))
  }

RunMultipleGenes <- function(genelist = "") {
  n_missing_rna <- length(genelist[!genelist %in% impulse_obj@vecAllIDs])
  print(paste0("Removing ",n_missing_rna," genes not in RNA."))
  genelist <- genelist[genelist %in% impulse_obj@vecAllIDs]
  n_missing_peaks <- length(genelist[!genelist %in% ATAC_all_peaks_anno.df$geneId])
  print(paste0("Removing ",n_missing_peaks, " genes, no associated peaks."))
  genelist <- genelist[genelist %in% ATAC_all_peaks_anno.df$geneId]
  # Multicore implementation
  temp <- bplapply(genelist, FUN = compareGeneToPeaks)
  print("bplApply Done")
  results.df <- do.call("rbind", temp)
  results.df$peakID <- ifelse(is.na(results.df$spearman),NA,rownames(results.df))
  results.df$Symbol <- lookup(results.df$gene, key.match = as.data.frame(GenesUniverse[,c(1,4)]))
  return(results.df)
}

```

```{r}
library(readr)
GRN_Genes <- read.csv(file = "./Imports/GRN_Genes.csv", header = F)
colnames(GRN_Genes) <- c("Symbol","ENSEMBL")
GenesUniverse <- read_csv("../RNA-Seq/Imports/GenesUniverse.csv")
GRN_enhancers <- RunMultipleGenes(genelist = as.character(GRN_Genes$ENSEMBL))
All_Genes_1 <- RunMultipleGenes(genelist = GenesUniverse$Chick_ENSEMBL[1:10000])
All_Genes_2 <- RunMultipleGenes(genelist = GenesUniverse$Chick_ENSEMBL[10001:24356])

All_Genes_Tests <- rbind(All_Genes, All_Genes_2)

# Remove rows with NA PeakIDs.

All_Genes_Tests <- All_Genes_Tests[!is.na(All_Genes_Tests$peakID),]

# look at the correlation of Promoter to Genes.




boxplot(All_Genes_Tests[All_Genes_Tests$annotation == "Distal Intergenic","spearman"])



lsgplotsGenes1 <- plotGenes(
  vecGeneIDs       = "ENSGALG00000012775",
  scaNTopIDs       = NULL,
  objectImpulseDE2 = impulse_obj,
  boolCaseCtrl     = F,
  boolSimplePlot   = T,
  dirOut           = NULL,
  strFileName      = NULL,
  vecRefPval       = NULL, 
  strNameRefMethod = NULL)

lsgplotsGenes2 <- plotGenes(
  vecGeneIDs       = "2-63870490-63870889",
  scaNTopIDs       = NULL,
  objectImpulseDE2 = impulse_obj2,
  boolCaseCtrl     = F,
  boolSimplePlot   = T,
  dirOut           = NULL,
  strFileName      = NULL,
  vecRefPval       = NULL, 
  strNameRefMethod = NULL)

cowplot::plot_grid(lsgplotsGenes1[[1]], lsgplotsGenes2[[1]], ncol = 1)

gene_fit <-
      getDfFit(
        vecGeneIDs = "ENSGALG00000000059",
        objectImpulseDE2 = impulse_obj,
        boolCaseCtrl = T,
        boolSimplePlot = T
      )

peak_fit <- 
  getDfFit(
        vecGeneIDs = "11-18842359-18842758",
        objectImpulseDE2 = impulse_obj2,
        boolCaseCtrl = F,
        boolSimplePlot = T
      )

cor(gene_fit$value[1:100], peak_fit$value, method = "spearman")
```
```{r}
compareGeneToPeaks(gene = "ENSGALG00000000059")
```

